---
title: "R Notebook"
output: html_notebook
---

```{r}
# install.packages("rstan", repos = "http://cran.rstudio.com/")
```

```{r}
# install.packages(c("devtools","mvtnorm","loo","coda", "dagitty"))
# library(devtools)
# devtools::install_github("rmcelreath/rethinking")

# install.packages("ggpubr")
```

```{r}
library(rethinking)
library(ggplot2)
library(ggpubr)

```

functions

calculate mean and std of data
```{r}
data_summary <- function(data, varname, groupnames){
  # Function to calculate the mean and the standard error
  # for each group
  #+++++++++++++++++++++++++
  # data : a data frame
  # varname : the name of a column containing the variable
  #to be summariezed
  # groupnames : vector of column names to be used as
  # grouping variables
  
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE)/sqrt(length(x[[col]])))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}
```

Function to simuate choice of n participant
```{r}
# simulate choices of n participants
simulate_choice <- function(n, trials, alpha, beta, slope) {
  
  # simulate all trials behavior for each sample of parameters (aka each imaginary participant)
  for (i in 1:n) {
    uL <- (risk - beta[i] * (ambig/2)) * value ^alpha[i] # subjective value of the lottery
    uR <- (risk.fixed - beta[i] * (ambig.fixed/2)) * value.fixed ^alpha[i] # subjective value of the reference
    prob <- 1 / (1 + exp(slope[i]*(uL-uR))) # choice probablity of the lottery
    
    choice <- numeric(0)
    choice <- c(choice, 1:nrow(trials))
    
    for (trial.idx in 1:nrow(trials)){
      choice[trial.idx] <- rbern(1, prob[trial.idx]) # choice sampled from bernoulli distribution
    }
    
    if (i == 1){
      simulation <- data.frame(id = rep(c(i), nrow(trials)), risk, ambig, value, risk.fixed, ambig.fixed, value.fixed, choice, alpha = rep(c(alpha[i]), nrow(trials)), beta = rep(c(beta[i]), nrow(trials)), slope = rep(c(slope[i]), nrow(trials)))
      
    } else {
      simulation <- rbind(simulation, data.frame(id = rep(c(i), nrow(trials)), risk, ambig, value, risk.fixed, ambig.fixed, value.fixed, choice, alpha = rep(c(alpha[i]), nrow(trials)), beta = rep(c(beta[i]), nrow(trials)), slope = rep(c(slope[i]), nrow(trials))))
    }
  }
  
  simulation$id <- as.factor(simulation$id)
  
  
  return(simulation)
}
```

Function to visualize individual subject choice data
```{r}
plot_choice <- function(simulation,n) {
  for (sub_idx in 1:n) {
    
    # single subject data
    simulation.sub <- simulation[simulation$id == sub_idx,]
    simulation.sub$value <- as.factor(simulation.sub$value)
    simulation.sub$risk <- as.factor(simulation.sub$risk)
    simulation.sub$ambig <- as.factor(simulation.sub$ambig)
    
    # monetary risky
    data_plot <- data_summary(simulation.sub[simulation.sub$ambig == 0,], varname = "choice", groupnames = c("value", "risk"))
    
    p1 <- ggplot(data_plot, aes(x=value, y=risk, fill=choice)) + 
      geom_tile(color = "white")+
      scale_x_discrete(breaks = value.uniq) +
      # scale_y_discrete(breaks = risk.uniq) +
      scale_fill_gradient(limits = c(0,1)) +
      theme(panel.grid = element_blank(),
            panel.background = element_blank()) +
      ggtitle(paste("Monetary Risky, Sub", as.character(sub_idx), ", alpha=", round(simulation.sub$alpha[1],2), ", slope=", round(simulation.sub$slope[1], 2))) +
      xlab("Value ($)") + ylab("Risk Level")
    
    # monetary ambiguous
    data_plot <- data_summary(simulation.sub[simulation.sub$ambig != 0,], varname = "choice", groupnames = c("value", "ambig"))
    
    p2 <- ggplot(data_plot, aes(x=value, y=ambig, fill=choice)) + 
      geom_tile(color = "white")+
      scale_x_discrete(breaks = value.uniq) +
      scale_y_discrete(breaks = ambig.uniq) +
      scale_fill_gradient(limits = c(0,1)) +
      theme(panel.grid = element_blank(),
            panel.background = element_blank()) +
      ggtitle(paste("Monetary Ambiguous, beta=", round(simulation.sub$beta[1],2))) +
      xlab("Value ($)") + ylab("Ambiguity Level")
    
    figure <- ggarrange(p1,p2,ncol = 1, nrow = 2) + border(size = 2)
    
    print(figure)
    
    # annotate_figure(figure,
    #                 top = text_grob(paste("Subject", sub_idx))
    #                 )
    
  }  
}

```

Function to plot one simulation (includeing CAPS)
```{r}
plot_choice_caps <- function(simulation, caps) {
  n = length(caps)
  for (sub_idx in 1:n) {
    
    # single subject data
    simulation.sub <- simulation[simulation$id == sub_idx,]
    simulation.sub$value <- as.factor(simulation.sub$value)
    simulation.sub$risk <- as.factor(simulation.sub$risk)
    simulation.sub$ambig <- as.factor(simulation.sub$ambig)
    
    # monetary risky
    data_plot <- data_summary(simulation.sub[simulation.sub$ambig == 0,], varname = "choice", groupnames = c("value", "risk"))
    
    p1 <- ggplot(data_plot, aes(x=value, y=risk, fill=choice)) + 
      geom_tile(color = "white")+
      scale_x_discrete(breaks = value.uniq) +
      # scale_y_discrete(breaks = risk.uniq) +
      scale_fill_gradient(limits = c(0,1)) +
      theme(panel.grid = element_blank(),
            panel.background = element_blank()) +
      ggtitle(paste("Monetary Risky, CAPS", round(caps[sub_idx],2), ", alpha=", round(simulation.sub$alpha[1],2), ", slope=", round(simulation.sub$slope[1], 2))) +
      xlab("Value ($)") + ylab("Risk Level")
    
    # monetary ambiguous
    data_plot <- data_summary(simulation.sub[simulation.sub$ambig != 0,], varname = "choice", groupnames = c("value", "ambig"))
    
    p2 <- ggplot(data_plot, aes(x=value, y=ambig, fill=choice)) + 
      geom_tile(color = "white")+
      scale_x_discrete(breaks = value.uniq) +
      scale_y_discrete(breaks = ambig.uniq) +
      scale_fill_gradient(limits = c(0,1)) +
      theme(panel.grid = element_blank(),
            panel.background = element_blank()) +
      ggtitle(paste("Monetary Ambiguous, beta=", round(simulation.sub$beta[1],2))) +
      xlab("Value ($)") + ylab("Ambiguity Level")
    
    figure <- ggarrange(p1,p2,ncol = 1, nrow = 2) + border(size = 2)
    
    print(figure)
    
    # annotate_figure(figure,
    #                 top = text_grob(paste("Subject", sub_idx))
    #                 )
    
  }  
}
```



make the task
```{r}
risk.uniq <- c(0.13, 0.25, 0.38, 0.5, 0.62, 0.75, 0.87)
ambig.uniq <- c(0.24, 0.5, 0.74, 1)

value.uniq <- c(5, 6, 7, 8, 10, 12, 14, 16, 19, 23, 27, 31, 37, 44, 52, 61, 73, 86, 101, 120)

rep.times <- 1

# make unique trials with combinations of uncertainty level and value
risk.level <- c(risk.uniq, rep(0.5, times = length(ambig.uniq))) 
ambig.level <- c(rep(0, times = length(risk.uniq)), ambig.uniq)

risk.trials <- expand.grid(risk.uniq, value.uniq)
colnames(risk.trials) <- c("risk", "value")
risk.trials$ambig <- 0

ambig.trials <- expand.grid(ambig.uniq, value.uniq)
colnames(ambig.trials) <- c("ambig", "value")
ambig.trials$risk <- 0.5

all.trials <- rbind(risk.trials, ambig.trials)

# fixed option
all.trials$risk.fixed <- 1
all.trials$ambig.fixed <- 0
all.trials$value.fixed <- 5

# repeat trials
trials <- data.frame(sapply(all.trials, rep.int, times=rep.times))

# number of trials
nrow(trials)
```



Define model
```{r}
risk <- trials$risk
ambig <- trials$ambig
value <- trials$value

risk.fixed <- trials$risk.fixed
ambig.fixed <- trials$ambig.fixed
value.fixed <- trials$value.fixed

m.choice <- alist(

  choice ~ dbern(1, prob), # sample from bernoulli distribution
  prob <- 1 / (1 + exp(slope*(uL-uR))), # choice probablity of the lottery
  uL <- (risk - beta * (ambig/2)) * v ^alpha, # subjective value of the lottery
  uR <- (risk.ref - beta * (ambig.ref/2)) * v.ref ^alpha, # subjective value of the reference

  # priors
  alpha ~ dnorm(1, 1),
  beta ~ dnorm(0, 1),
  slope ~ dnorm(-2, 3)
)

```

priors
```{r}
set.seed(2020)

# number of simulations
n <- 20

# sample parameters from prior
# risk attitude
alpha <- rnorm(n, 1, 1)
# ambiguity attitude
beta <- rnorm(n, 0, 1)
# inverse temperature
slope <- rnorm(n, -2, 3)

# plot
hist(alpha, nclass =10)
# dens(alpha, add = TRUE)
hist(beta, nclass =10)
hist(slope, nclass =10)
```




```{r}
simulation <- simulate_choice(n,trials,alpha,beta,slope)
View(simulation)

```


Plot each simulation's choice
```{r}
plot_choice(simulation,n)
```



Risk attitude affected by symptoms

Generate participants (CAPS, PTSD symptom)
```{r}
set.seed(2020)

# number of participants
n.p <- 30
# generate n participants with PTSD symptom severity
caps <- runif(n.p, min = 0, max = 100) # uniform distribution, not really in the real data
# standardize
caps.z <- scale(caps)
hist(caps, nclass = 10)
```

Priors
```{r}
# number of simulation
n.s <- 10

# sample parameters from prior

# risk attitude, affected by CAPS (linear)
# m <- rnorm(n.s, 1, 1)
# k <- rnorm(n.s, 0, 1)
# sigma <- runif(n.s, 0, 15)

m <- rnorm(n.s, 0.5, 0.2)
k <- -rlnorm(n.s, 0, 0.5)

sigma <- runif(n.s, 0, 2)

alpha <- matrix(, nrow = n.s, ncol = n.p)
for (i in 1:n.s){ # number of simulation          
  miu <- m[i] + k[i]*caps.z
  for (j in 1:n.p){ # number of participants in each simulation
    alpha[i,j] <- rnorm(1, miu[j], sigma[i]) 
  }
}

# ambiguity attitude
beta <- matrix(, nrow = n.s, ncol = n.p)
for (i in 1:n.s) {
  beta[i, ] <- rnorm(n.p, 0, 1)
}


# inverse temperature
slope <- matrix(, nrow = n.s, ncol = n.p)
for (i in 1:n.s) {
  slope[i, ] <- rnorm(n.p, -3, 1)
}

# to summarize, there are 5 parameters to generate the choice behavior, and each is assumed a prior
  # m (average of risk attitude), k(linear slope of how CAPS influence alpha), sigma(variace of alpha's)
# beta (ambiguity attitude), slope (inverse temperature)
```

Plot priors of k (linear slope of how CAPS influence alpha) and m (average of risk attitude)
```{r}
hist(k)
hist(m)
```

Plot all simulation
Risky attitude and caps
```{r}
plot(caps, alpha[1,], type = "n",  ylim=c(0, 2), xlim=c(0, 100),ylab = "alpha")
for (i in 1:n.s){
  lines(caps, k[i]*caps.z+m[i], col = col.alpha("black", 0.1))
}
```

Look at the first simulation
simulate choice
```{r}
sim_idx = 5
# hist(alpha[sim_idx,])
# hist(beta[sim_idx,])
# hist(slope[sim_idx,])
# 
simulation.caps <- simulate_choice(n.p, trials, alpha[sim_idx,], beta[sim_idx,], slope[sim_idx,])

# View(simulation.caps)
```

Plot the risk attitudes and caps
```{r}
participant <- data.frame(caps,alpha = alpha[sim_idx,], beta = beta[sim_idx,], slope = slope[sim_idx,], k = k[sim_idx], m = m[sim_idx], sigma = sigma[sim_idx])

ggplot(data = participant, aes(x=caps, y=alpha)) +
  geom_point() +
  geom_line(aes(caps, k[1]*caps.z + m[1])) +
  ggtitle(paste("Risk attitude, alpha =", round(k[1],2), "* zscored caps +", round(m[1],2) ))

ggplot(data = participant, aes(x=caps, y=beta)) +
  geom_point() +
  ggtitle(("Ambiguity attitude"))

ggplot(data = participant, aes(x=caps, y=slope)) +
  geom_point() +
  ggtitle(("Inverse temperature"))
```



```{r}
plot_choice_caps(simulation.caps, caps)
```


Calculate choice (model free attitude)
```{r}

for (i in 1:n.p){
  participant$risk_choice[i] = mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig==0])
  
participant$ambig_choice_corr[i] = mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig!=0]) - mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig==0 & simulation.caps$risk==0.5])
}

```

Plot caps and choice
```{r}
ggplot(participant, aes(x=caps, y=risk_choice)) +
  geom_point() +
  ggtitle("Choice of risky trials")

ggplot(participant, aes(x=caps, y=ambig_choice_corr)) +
  geom_point() +
  ggtitle("Choice of Ambiguous trials (Corrected)")
```

Look at 5 simulations
```{r}
# plot model risk attitude and caps
plot(caps, alpha[1,], type = "n",  ylim=c(0, 2), xlim=c(0, 100),ylab = "alpha")
for (i in 1:5){
  lines(caps, k[i]*caps.z+m[i], col = col.alpha("black", 0.1))
}

# plot choice behavior and caps
for (sim_idx in 1:5) {
  simulation.caps <- simulate_choice(n.p, trials, alpha[sim_idx,], beta[sim_idx,], slope[sim_idx,])
  
  participant <- data.frame(caps,alpha = alpha[sim_idx,], beta = beta[sim_idx,], slope = slope[sim_idx,], k = k[sim_idx], m = m[sim_idx], sigma = sigma[sim_idx])
  
  
  for (i in 1:n.p){
    participant$risk_choice[i] = mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig==0])
    
  participant$ambig_choice_corr[i] = mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig!=0]) - mean(simulation.caps$choice[simulation.caps$id==i & simulation.caps$ambig==0 & simulation.caps$risk==0.5])
  }
  
  f <- ggplot(participant, aes(x=caps, y=risk_choice)) +
    geom_point() +
    ggtitle("Choice of risky trials")  
  
  print(f)
}


```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

