---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("rstan", repos = "http://cran.rstudio.com/")
```

```{r}
install.packages(c("devtools","mvtnorm","loo","coda", "dagitty"))
library(devtools)
devtools::install_github("rmcelreath/rethinking")
```

```{r}
library(rethinking)
```

load data
```{r}
data("Howell1")
d <- Howell1
```

Problem 1
fit model
```{r}
d <- Howell1[Howell1$age > 18,]

xbar <- mean(d$weight)

# specify model
m1 <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b*(weight - xbar),
  a ~ dnorm(178, 20),
  b ~ dlnorm(0, 1),
  sigma ~ dunif(0, 50)
)

# fit
m1.f <- quap(m1, data = d)
```

sample posterior
```{r}
# sample parameters from posterior
post <- extract.samples(m1.f, n = 10000)
str(post)

# weight to predict
predict.x <- c(45, 40, 65, 31, 53)
data.predict <- data.frame(x = predict.x)
View(data.predict)

# sample y from posterior
mu <- matrix(NA, nrow = 10000, ncol = length(data.predict$x))
y <- matrix(NA, nrow = 10000, ncol = length(data.predict$x))

for (i in 1:10000) {
  mu[i, ] <- post$a[i] + post$b[i] * (data.predict$x - xbar)
  for (j in 1:length(data.predict$x)) {
    y[i, j] <- rnorm(1, mu[i, j], post$sigma[i])
  }
}

# or use the sim() function
# y <- sim(m1.f, data = data.frame(weight = predict.x))
str(y)

# plot
for (i in 1:length(data.predict$x)) {
  plot(density(y[, i]), col = "blue")
}  

# mean and compatibility interval  
# mean
data.predict$y <- colMeans(y)
data.predict$y

# 89% compatibility interval
ci_89 <- apply(y, 2, quantile, c(0.055, 0.945))
# or use the PI function
# ci_89 <- apply(y, 2, PI, 0.89)
ci_89

data.predict$ci_89l <- ci_89[1,]
data.predict$ci_89u <- ci_89[2,]

data.predict
```

Problem 2

Visualize
```{r}
d <- Howell1
d$weight.log <- log(d$weight)

# plot, looks linear?
plot(d$weight.log, d$height, col = col.alpha(rangi2,0.5))
plot(d$weight, d$height, col = col.alpha(rangi2,0.5))
```

fit model
```{r}
xbar <- mean(d$weight.log)

m2 <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b*(weight.log - xbar),
  a ~ dnorm(178, 20),
  b ~ dlnorm(0, 1),
  sigma ~ dunif(0, 50)
)

m2.f <- quap(m2, data = d)
```

plot posterior prediction
```{r}
weight.predict <- seq(5, 65, 2.5)
weight.log.predict <- log(seq(5, 65, 2.5))
height.predict <- sim(m2.f, data = data.frame(weight.log = weight.log.predict))

# look at size
str(height.predict)

# plot predicted height
plot(d$weight.log, d$height, type = "n")

for (i in 1:100) {
  points(weight.log.predict, height.predict[i, ], pch=16, col = col.alpha(rangi2,0.1))
}

# plot raw data and fitted line with 89% interval
height.predict.mean <- apply(height.predict, 2, mean)
height.predict.PI <- apply(height.predict, 2, PI, prob = 0.89)
plot(d$weight.log, d$height, col = col.alpha(rangi2,0.5))
lines(weight.log.predict, height.predict.mean)
shade(height.predict.PI, weight.log.predict)

```

plot posterior prediction, (weight instead of weight.log on x axis)
```{r}
weight.predict <- seq(5, 65, 2.5)
weight.log.predict <- log(seq(5, 65, 2.5))
# weight.log.predict <-d$weight.log
height.predict <- sim(m2.f, data = data.frame(weight.log = weight.log.predict))

# look at size
str(height.predict)

# plot predicted height
plot(d$weight, d$height, type = "n")

for (i in 1:100) {
  points(weight.predict, height.predict[i, ], pch=16, col = col.alpha(rangi2,0.1))
}

# plot raw data and fitted line with 89% interval
height.predict.mean <- apply(height.predict, 2, mean)
height.predict.PI <- apply(height.predict, 2, PI, prob = 0.89)
plot(d$weight, d$height, col = col.alpha(rangi2,0.5))
lines(weight.predict, height.predict.mean)
shade(height.predict.PI, weight.predict)

```
Problem 3

fit parabola model
```{r}
d$weight_s <- (d$weight - mean(d$weight))/sd(d$weight)
d$weight_s2 <- d$weight_s^2

m.parab <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b1*weight_s + b2*weight_s2,
  a ~ dnorm(178, 20),
  b1 ~ dlnorm(0, 1),
  b2 ~ dlnorm(0, 1),
  sigma ~ dunif(0, 50) 
)

# fit
m.parab.f <- quap(m.parab, data = d)
```

simulate, very random priors
```{r}
set.seed(2020)

# mean of the sample weight
xbar <- mean(d$weight)

# min(d$weight)
# max(d$weight)
# mean(d$height)

# pick weight to predict
weight.predict <- seq(5, 65, 2.5)
weight.predict

# number of samples
n <- 200

# sample parameters from prior
# very uninformative prior
a <- rnorm(n, 178, 20)
b1 <- rnorm(n, 0, 1)
b2 <- rnorm(n, 0, 1)

plot(d$weight, d$height, type = "n", ylim=c(50, 240), xlim=c(0, 65), yaxt="n")
# plot(d$weight, d$height, col=col.alpha(rangi2, 0.3), ylim=c(50, 240), xlim=c(0, 65), yaxt="n")

axis(side=2, at=seq(60, 240, 20) )

for (i in 1:n) {
  mu <- a[i] + b1[i]*(weight.predict - xbar) + b2[i]*(weight.predict - xbar)^2
  lines(weight.predict, mu, col = col.alpha("black", 0.1))
}
```

simulate, tweaking priors
```{r}
set.seed(2020)

# mean of the sample weight
xbar <- mean(d$weight)

# min(d$weight)
# max(d$weight)
# mean(d$height)

# pick weight to predict
weight.predict <- seq(5, 65, 2.5)

# number of samples
n <- 200

# sample parameters from prior
# try differet priors
a <- rnorm(n, 150, 10)
# b1 <- rlnorm(n, 0, 1)
# b2 <- rlnorm(n, 0, 1)
b1 <- rnorm(n, 2, 0.05)
b2 <- rnorm(n, -0.05, 0.0005)

# plot(d$weight, d$height, type = "n", ylim=c(50, 240), xlim=c(0, 65), yaxt="n")
plot(d$weight, d$height, col=col.alpha(rangi2, 0.3), ylim=c(50, 240), xlim=c(0, 65), yaxt="n")

axis(side=2, at=seq(60, 240, 20) )

for (i in 1:n) {
  mu <- a[i] + b1[i]*(weight.predict - xbar) + b2[i]*(weight.predict - xbar)^2
  lines(weight.predict, mu, col = col.alpha("black", 0.1))
}
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

